---
title: "Networked Data Lab, hospital use among CEV people during COVID-19"
author: "Sebastien Peytrignet"
date: "4/14/2021"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    toc: TRUE
---

```{r setup.A, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Load packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,stringr,sp,ggplot2,readODS,
               gmodels,data.table,tibble,tidyverse,
               readxl,plotly,hrbrthemes,RColorBrewer,
               knitr,kableExtra)

rm(list = ls())

#Directories
rawdatadir <- "M:/Analytics/Networked Data Lab/Partner outputs/Output 3 HC usage/"
gitdir <- "M:/Analytics/Sebastien/GitHub/NDL_Output3_Hospital_care_CEV/"
```

## Reading in data

```{r setup.B, echo=TRUE}
#Function to clean column names
clean_colnames <- function(dataset){
  names(dataset) <- str_replace_all(names(dataset),"\\.","_")
  return(dataset)
}

#Files directory
setwd(gitdir)

##################################################
################# Auxiliary data #################
##################################################

#Ranking of types of service
bl.orders <- data.frame(order=1:4,breakdown.level=c("outpatient_attendance","AE_attendance","admissions_electives","admissions_emergency"))

#Order of months
month.lookup <- data.frame(month.number=c("01","02","03","04","05","06","07","08","09","10","11","12"),month.abbr=month.abb,month.full=month.name)

##########################################################
################# Service use statistics #################
##########################################################

table1.all.partners <- fread("Visualize data/Data/table1.all.partners.csv",
                             header=TRUE, sep=",", check.names=T) %>%
  left_join(.,bl.orders,by="breakdown.level") %>%
  mutate(.,year=word(year_month, 1, sep = fixed('-')),month=word(year_month, 2, sep = fixed('-'))) %>%
  left_join(.,month.lookup,by=c("month"="month.number")) %>%
  mutate(.,date_text=paste(month.full,year,sep=" "),
         year_month_day=paste(year_month,"01",sep="-")) %>%
  filter(.,!(breakdown.level %in% c("admissions_other")))

#Changes in service use data
#This step creates a new column for each row of data (activity levels one year ago)
#and computes the absolute and relative difference

table1.all.partners.pcts <-  table1.all.partners %>%
  filter(., !(strata %in% c("unknown","unknown/other"))) %>%
  select(.,partner,year_month,year_month_day,strata,breakdown,breakdown.level,
         number.events.sdc,total.patients,year,month,date_text) %>%
  mutate(.,year_month_day=lubridate::ymd(year_month_day),
         year_l1=(as.numeric(year)-1)) %>%
  mutate(.,year_month_day_l1=lubridate::ymd(paste(year_l1,month,"01",sep="-")),
         keyone=paste(partner,breakdown,strata,breakdown.level,sep="-"))

#Lags data
table1.all.partners.pcts.l1 <- table1.all.partners.pcts %>%
  select(.,keyone,year_month_day,number.events.sdc) %>%
  dplyr::rename(.,number.events.sdc_l1=number.events.sdc,
                year_month_day_l1=year_month_day)

#Merged back into main data
table1.all.partners.pcts <- left_join(table1.all.partners.pcts,
                                      table1.all.partners.pcts.l1,
                                      by=c("keyone","year_month_day_l1")) %>%
  select(.,-c("keyone","year_l1")) %>%
  mutate(.,pct_change_20to19=(number.events.sdc-number.events.sdc_l1)/number.events.sdc_l1*100,
         abs_change_12m=(number.events.sdc-number.events.sdc_l1)) %>%
  filter(.,year_month>="2019-07") %>%
  mutate(.,year_month_day=lubridate::ymd(year_month_day)) %>%
  arrange(.,partner,breakdown,strata,breakdown.level,year_month)
rm(table1.all.partners.pcts.l1)

####################################################
################# COVID admissions #################
####################################################

table2.all.partners <- fread("Visualize data/Data/table2.all.partners.csv",
                             header=TRUE, sep=",", check.names=T)

##########################################
################# Deaths #################
##########################################

table3.all.partners <- fread("Visualize data/Data/table3.all.partners.csv",
                             header=TRUE, sep=",", check.names=T)

###############################################################################################
################# Sup Table 1: Admissions by ICD-10 chapter of main diagnosis #################
###############################################################################################

tableS1.all.partners <- fread("Visualize data/Data/tableS1.all.partners.csv",
                             header=TRUE, sep=",", check.names=T)

###############################################################################
################# Sup Table 2: General population comparisons #################
###############################################################################

#Comparison to general population using data from North West London
tableS2.all.partners <- fread("Visualize data/Data/tableS2.all.partners.csv",
                             header=TRUE, sep=",", check.names=T) %>%
  filter(.,partner!="Liverpool-Wirral")

#Add % change data

tableS2.all.partners.pcts <-  tableS2.all.partners %>%
  mutate(.,year_month_day=lubridate::ymd(year_month_day),
         year_l1=(as.numeric(year)-1)) %>%
  mutate(.,year_month_day_l1=lubridate::ymd(paste(year_l1,month,"01",sep="-")),
         keyone=paste(partner,cohort,breakdown.level,sep="-"))

tableS2.all.partners.pcts.l1 <- tableS2.all.partners.pcts %>%
  select(.,keyone,year_month_day,number.events.sdc) %>%
  dplyr::rename(.,number.events.sdc_l1=number.events.sdc,
                year_month_day_l1=year_month_day)

tableS2.all.partners.pcts <- left_join(tableS2.all.partners.pcts,
                                       tableS2.all.partners.pcts.l1,
                                      by=c("keyone","year_month_day_l1")) %>%
  select(.,-c("keyone","year_l1")) %>%
  mutate(.,pct_change_20to19=(number.events.sdc-number.events.sdc_l1)/number.events.sdc_l1*100,
         abs_change_12m=(number.events.sdc-number.events.sdc_l1)) %>%
  mutate(.,year_month_day=lubridate::ymd(year_month_day)) %>%
  filter(.,year_month>="2019-07") %>%
  arrange(.,partner,cohort,breakdown.level,year_month)
rm(tableS2.all.partners.pcts.l1)
```

## % reduction in activity in April 2020 vs. April 2019 (elective admissions, emergency outpatient, A&E and outpatient)

### Pooled

```{r apr.pct.change.pooled, echo=TRUE, message=FALSE, warning=FALSE}

pooled.drop.april <- table1.all.partners.pcts %>% 
  filter(.,strata=="overall",year_month=="2020-04",
         breakdown.level!="admissions_all") %>%
  select(.,breakdown.level,partner,year_month,
         number.events.sdc_l1,number.events.sdc,
         abs_change_12m) %>%
  group_by(breakdown.level) %>%
  summarise(.,number.events.sdc_l1=sum(number.events.sdc_l1,na.rm=TRUE),
            abs_change_12m=sum(abs_change_12m,na.rm=TRUE)) %>%
  ungroup() %>%
  mutate(.,pct_change=abs_change_12m/number.events.sdc_l1*100) %>%
  select(.,breakdown.level,abs_change_12m,pct_change) %>%
  clean_colnames(.)

pooled.drop.april %>%
  kable(.,digits=1,format.args = list(big.mark = ",")) %>%
  kable_styling()
```

### By partner

```{r apr.pct.change, echo=TRUE, message=FALSE, warning=FALSE}

drop.april <- table1.all.partners.pcts %>%
  filter(.,strata=="overall",year_month=="2020-04",
         breakdown.level!="admissions_all") %>%
  select(.,breakdown.level,partner,year_month,
         number.events.sdc_l1,number.events.sdc,abs_change_12m) %>%
  group_by(breakdown.level,partner) %>%
  summarise(.,number.events.sdc_l1=sum(number.events.sdc_l1,na.rm=TRUE),
            abs_change_12m=sum(abs_change_12m,na.rm=TRUE)) %>%
  ungroup() %>%
  mutate(.,pct_change=abs_change_12m/number.events.sdc_l1*100) %>%
  select(.,breakdown.level,partner,abs_change_12m,pct_change) %>%
  arrange(.,breakdown.level,desc(pct_change)) %>%
  clean_colnames(.)

drop.april %>%
  kable(.,digits=1,format.args = list(big.mark = ",")) %>%
  kable_styling(., full_width = F, font_size= 8)
```

## % of admissions that were COVID-related

### Pooled

```{r pct.covid.pooled, echo=TRUE, message=FALSE, warning=FALSE}

covid.adm.data <- table1.all.partners %>%
  filter(.,strata=="overall",breakdown.level=="admissions_all",
    year_month>="2020-03") %>%
  select(.,breakdown.level,partner,year_month,number.patients.sdc,
         number.patients.covid,total.patients)

#Rate of admitted patients who 'had COVID' (pooled)
pooled.covid  <- covid.adm.data %>%
  summarise(across(c("number.patients.sdc", "number.patients.covid"), ~ sum(.x, na.rm = TRUE))) %>%
  mutate(.,rate_adm_covid=number.patients.covid/number.patients.sdc*100) %>%
  clean_colnames(.)

pooled.covid %>%
  kable(.,digits=1,format.args = list(big.mark = ",")) %>%
  kable_styling(., full_width = F)

```

### By partner

```{r pct.covid.partners, echo=TRUE, message=FALSE, warning=FALSE}

covid.adm.data <- table1.all.partners %>%
  filter(.,strata=="overall",breakdown.level=="admissions_all",
    year_month>="2020-03") %>%
  select(.,breakdown.level,partner,year_month,number.patients.sdc,
         number.patients.covid,total.patients)

#Rate of admitted patients who 'had COVID' (unpooled)
allp.covid <- covid.adm.data %>%
  group_by(partner) %>%
  summarise(across(c("number.patients.sdc", "number.patients.covid"), ~ sum(.x, na.rm = TRUE))) %>%
  mutate(.,rate_adm_covid=number.patients.covid/number.patients.sdc*100) %>%
  ungroup() %>%
  arrange(.,rate_adm_covid) %>%
  clean_colnames(.)

allp.covid %>%
  kable(.,digits=1,format.args = list(big.mark = ",")) %>%
  kable_styling(., full_width = F)

```

## Total number of patients with COVID

```{r total.covid, echo=TRUE, message=FALSE, warning=FALSE}

total.covid.patients <- table1.all.partners %>% filter(.,strata=="overall",
                         breakdown.level=="admissions_all",
                         year_month>="2020-03") %>%
  select(.,breakdown.level,partner,year_month,number.patients.sdc,
         number.patients.covid,total.patients) %>%
  group_by(breakdown.level) %>%
  summarise(.,number.patients.covid=sum(number.patients.covid,na.rm=TRUE)) %>%
  ungroup() %>%
  clean_colnames(.)

total.covid.patients %>%
  kable(.,digits=1,format.args = list(big.mark = ",")) %>%
  kable_styling(., full_width = F)

```

## Total number of deaths

```{r total.deaths, echo=TRUE, message=FALSE, warning=FALSE}

total.deaths <- table3.all.partners %>%
  filter(.,location=="anywhere",breakdown.level=="overall") %>%
  select(.,number.patients.sdc) %>%
  summarise(total.deaths=sum(number.patients.sdc,na.rm=TRUE)) %>%
  clean_colnames(.)

total.deaths %>%
  kable(.,digits=1,format.args = list(big.mark = ",")) %>%
  kable_styling(., full_width = F)

```

## Cumulative mortality rate

### Pooled

```{r pooled.cum.deaths, echo=TRUE, message=FALSE, warning=FALSE}

total.deaths.cum.pooled <- table3.all.partners %>%
  filter(.,location=="anywhere",breakdown.level=="overall") %>%
  select(.,location,year_month,partner,number.patients.sdc,total.patients) %>%
  group_by(partner) %>%
  summarise(.,number.patients.sdc=sum(number.patients.sdc,na.rm=TRUE),
            total.patients=first(total.patients)) %>%
  ungroup() %>%
  summarise(.,number.patients.sdc=sum(number.patients.sdc,na.rm=TRUE),
            total.patients=sum(total.patients,na.rm=TRUE)) %>%
  mutate(.,cum_rate=number.patients.sdc/total.patients*100) %>%
  clean_colnames(.)

total.deaths.cum.pooled %>%
  kable(.,digits=1,format.args = list(big.mark = ",")) %>%
  kable_styling(., full_width = F)

```

### All partners

```{r cum.deaths, echo=TRUE, message=FALSE, warning=FALSE}

total.deaths.cum <- table3.all.partners %>% 
  filter(.,year_month=="2020-07",location=="anywhere",breakdown.level=="overall") %>%
  select(.,partner,death_cum_pct) %>%
  arrange(.,death_cum_pct) %>%
  clean_colnames(.)

total.deaths.cum %>%
  kable(.,digits=1,format.args = list(big.mark = ",")) %>%
  kable_styling(., full_width = F)

```

## Changes by ICD-10 chapter {.tabset}

### Elective admissions

```{r, abs.elective, echo=FALSE, message=FALSE, warning=FALSE,fig.width=10,fig.height=10}

tableS1.all.partners %>%
  filter(.,!(ICD10.chapter %in% c("No match with ICD-10 chapters","22 - Codes for special purposes"))) %>%
  filter(.,type=="elective",partner=="Average") %>%
  select(.,partner,ICD10.chapter,number.events.2019.sdc,number.events.2020.sdc) %>%
  mutate(.,abs_change_2019_to_2020=(number.events.2020.sdc-number.events.2019.sdc),
         pct_change_2019_to_2020=round((number.events.2020.sdc-number.events.2019.sdc)/number.events.2019.sdc*100,1)) %>%
  arrange(.,-number.events.2020.sdc) %>%
  kable(.,format.args = list(big.mark = ",")) %>%
  kable_styling()

```

### Emergency admissions

```{r, abs.emergency, echo=FALSE, message=FALSE, warning=FALSE,fig.width=10,fig.height=10}

tableS1.all.partners %>%
  filter(.,!(ICD10.chapter %in% c("No match with ICD-10 chapters","22 - Codes for special purposes"))) %>%
  filter(.,type=="emergency",partner=="Average") %>%
  select(.,partner,ICD10.chapter,number.events.2019.sdc,number.events.2020.sdc) %>%
  mutate(.,abs_change_2019_to_2020=(number.events.2020.sdc-number.events.2019.sdc),
         pct_change_2019_to_2020=round((number.events.2020.sdc-number.events.2019.sdc)/number.events.2019.sdc*100,1)) %>%
  arrange(.,-number.events.2020.sdc) %>%
  kable(.,format.args = list(big.mark = ",")) %>%
  kable_styling()

```

## General population comparison in North West London

```{r, nwl.genpop, echo=FALSE, message=FALSE, warning=FALSE,fig.width=10,fig.height=10}
tableS2.all.partners %>%
  filter(.,!(breakdown.level %in% c("admissions_all","admissions_other"))) %>%
  filter(.,(year_month=="2020-02"|year_month=="2020-04")) %>%
  mutate(.,year_month=ifelse(year_month=="2020-02","feb20","apr20")) %>%
  select(.,breakdown.level,partner,cohort,year_month,number.events.sdc,total.patients,cases.per.100) %>%
  pivot_wider(.,names_from = year_month,
              names_sep = ".",
              values_from = c(number.events.sdc,cases.per.100,total.patients)) %>%
  mutate(.,pct_change=(number.events.sdc.apr20-number.events.sdc.feb20)/number.events.sdc.feb20*100) %>%
  arrange(.,breakdown.level,cohort) %>% 
kable(.,digits=1,format.args = list(big.mark = ",")) %>%
  kable_styling(., full_width = F)
```
