
# Libraries
```{r, message = F}
library(tidyverse)

.libPaths(
  c('H:/DaSH416/R_libraries/R_Transfer_25062020',
    'H:/DaSH416/R_libraries/R_Libraries_15072020',
    'H:/DaSH416/R_libraries/R_Transfer_06012021'))

library(here)
library(vroom)
library(lubridate)
library(janitor)

options(scipen = 999)
```

# Shielding list data
```{r message = F}
#tidy and combine "pregnant" with "other" group

shielding_long <-
  vroom(here("original_data", 'Dash416_Shield20201012_Release.csv'),
        delim = "¬") %>%
  clean_names() %>%
  rename(id = dash416_release_uid) %>%
  mutate(date_addition = as_date(ymd_hms(earliest_addition_this_chi)),
         date_death = as_date(ymd(trak_person_deceased_date))) %>%
  select(id,
         group,
         origin_for_this_group,
         date_addition,
         removal_description,
         date_death) %>%
  mutate(group = ifelse(group == 6, 7, group)) %>%
  distinct()

#add translation of shielding group codes
shielding_long <-
  vroom(here("original_data", "shielding_codes_short.csv"),
        delim = ",") %>%
  left_join(shielding_long, ., by = c("group" = "shielding_group"))

#demographic data
demog <-
  vroom(here("original_data", 'Dash416_Shield20201012_Demographic_Release.csv'),
        delim = "¬") %>%
  rename(id = Dash416_Release_UID) %>%
  clean_names() %>%
  select(id, calc_sex, calc_dob) 

#small area statistics about home 
vulnerability <-
  vroom(here("original_data", 'Dash416_Shield20201012_vulnerability_Release.csv'),
        delim = "¬") %>%
  clean_names() %>%
  rename(id = dash416_release_uid,
         ur_name = u_rname) %>%
  mutate(simd_quintile = as_factor(simd2020v2_quintile)) %>%
  select(id, simd_quintile, ur_name)
```


# Population exclusions
```{r}
#anyone not on shielding list before 1 Aug
#anyone who died before 28 March
#false positives
#moved from Scotland
shielding_long <-
  shielding_long %>%
  filter(date_addition <= ymd("2020-07-31")) %>%
  filter(date_death > ymd("2020-03-28") |
           is.na(date_death)) %>%
  filter(
    !removal_description %in%
      c("FalsePosLungCancer", "False Positive", "Moved out of Scotland")) 

unique_ids <- unique(shielding_long$id)
n_shielders = length(unique_ids)
```


# Tidy demographic data
```{r}
#keep only people in NDL cohort
demog <- 
  demog %>%
    filter(id %in% unique_ids)

#create variables for sex and age band
demog <- demog %>%
  mutate(
    sex = if_else(calc_sex == 1, "F", "M"),
    dob = ymd(calc_dob),
    age = as.period(interval(dob, ymd("2020-03-01")), units = "years")$year,
    age_band = 
      case_when(
        age %in% c(0:29) ~ "0-29",
        age %in% c(30:49) ~"30-49",
        age %in% c(50:69) ~"50-69",
        age %in% c(70:120) ~"70+")) %>%
  select(-calc_sex, -calc_dob)

#create age band variable
demog$age_band <- factor(demog$age_band, levels = c("0-29", "30-49", "50-69", "70+"))

#add details of shielding reasons to demographics
demog <- 
  demog %>%
  left_join(., vulnerability, by = "id")

#add deaths to demog
demog <-
  shielding_long %>%
  select(id, date_death) %>%
  unique() %>%
  left_join(demog, ., by = "id") 

#add death month field
#NB exclude deaths after July for NDL
demog <-
  demog %>%
  mutate(month_death = floor_date(date_death, "month"),
         month_death_ndl = 
           as_date(ifelse(
             month_death > ymd("2020-07-31"), NA, month_death)))


#add some demographics values to long shielding list
shielding_long <-
  demog %>%
    select(id, age, age_band, sex, simd_quintile, ur_name) %>%
  left_join(shielding_long, ., by = "id") 

rm(vulnerability)

#count thow many have SIMD data
n_simd = nrow(filter(demog, !is.na(simd_quintile)))
```

#Load and tidy A&E data
```{r}
ae <-
  vroom(here("original_data", 'Dash416_HI_ED_Attendances_Release_v1.csv'),
        delim = "¬") %>%
  clean_names() %>%
  rename(id = dash416_release_uid,
         date = arrival_date) %>%
  select(id, date)

#filter for NDL date range and cohort
#add month variable
ae <- 
  filter(ae, 
         date >= ymd("2018-03-01"),
         date <= ymd("2020-07-31"),
         id %in% unique_ids) %>%
  mutate(date_month = as_date(floor_date(date, "month")))

#add demog data to A&E
ae <-
  ae %>%
  left_join(., demog, by = "id")
```

#A&E attendances tables
```{r}
ae %>%
  group_by(date_month, .drop = F) %>%
  summarise(ae_attendances = n(),
            people_attending = n_distinct(id),
            cohort_size = n_shielders) %>%
  write_csv(., "ae_attendances_total.csv")

ae %>%
  group_by(date_month, age_band, .drop = F) %>%
  summarise(ae_attendances = n(),
            people_attending = n_distinct(id),
            cohort_size = n_shielders) %>%
  write_csv(., "ae_attendances_by_age.csv")

ae %>%
  group_by(date_month, sex, .drop = F) %>%
  summarise(ae_attendances = n(),
            people_attending = n_distinct(id),
            cohort_size = n_shielders) %>%
  write_csv(., "ae_attendances_by_sex.csv")

#remove NAs due to small numbers
ae %>%
  filter(!is.na(simd_quintile)) %>%
  group_by(date_month, simd_quintile, .drop = F) %>%
  summarise(ae_attendances = n(),
            people_attending = n_distinct(id),
            cohort_size = n_simd) %>%
  write_csv(., "ae_attendances_by_deprivation.csv")
```

#Load and tidy outpatient attendances
```{r}
smr00 <-
  vroom(here("original_data", 'Dash416_Shield20201012_SMR00_Release.csv'),
        delim = ",") %>%
  clean_names() %>%
  rename(id = dash416_release_uid,
         date_clinic = clinic_date) %>%
  mutate(date_clinic = as_date(date_clinic)) %>%
  select(id, date_clinic, clinic_attendance, specialty)

#select date range, attenders, and those in NDL population
smr00 <- 
  filter(smr00, 
         date_clinic >= ymd("2018-03-01"),
         date_clinic <= ymd("2020-07-31"),
         id %in% unique_ids,
         clinic_attendance == 1) %>%
  select(-clinic_attendance) %>%
  mutate(date_month = floor_date(date_clinic, "month"))

#add demog data to hospital admissions
smr00 <-
  smr00 %>%
  left_join(., demog, by = "id") 
```

#Outpatient attendances tables
```{r}
smr00 %>%
  group_by(date_month, .drop = F) %>%
  summarise(outpatient_attendances = n(),
            people_attending = n_distinct(id),
            cohort_size = n_shielders) %>%
  write_csv(., "outpatient_attendances_total.csv")

smr00 %>%
  group_by(date_month, age_band, .drop = F) %>%
  summarise(outpatient_attendances = n(),
            people_attending = n_distinct(id),
            cohort_size = n_shielders) %>%
  write_csv(., "outpatient_attendances_by_age.csv")

smr00 %>%
  group_by(date_month, sex, .drop = F) %>%
  summarise(outpatient_attendances = n(),
            people_attending = n_distinct(id),
            cohort_size = n_shielders) %>%
  write_csv(., "outpatient_attendances_by_sex.csv")

#remove NAs due to small numbers
smr00 %>%
  filter(!is.na(simd_quintile)) %>%
  group_by(date_month, simd_quintile, .drop = F) %>%
  summarise(outpatient_attendances = n(),
            people_attending = n_distinct(id),
            cohort_size = n_simd) %>%
  write_csv(., "outpatient_attendances_by_deprivation.csv")
```


# Load and tidy hospital admissions
```{r}
#load data, rename variables, format dates, add episode ID and date month, select only variables needed

smr01_long <-
  vroom(here("original_data", 'Dash416_Shield20201012_SMR01_Release.csv'),
        delim = "¬") %>%
  clean_names() %>%
  rename(id = dash416_release_uid,
         main_condition = main_cod,
         other_condition1 = oc1,
         other_condition2 = oc2,
         other_condition3 = oc3,
         other_condition4 = oc4,
         other_condition5 = oc5) %>%
  mutate(
    date_episode_start = as_date(ymd(date_episode_start)),
    date_episode_end = as_date(ymd(date_episode_end)),
    date_admission = as_date(ymd(adm_date)),
    date_discharge = as_date(ymd(date_discharge)),
    episode_id = paste(id, date_episode_start, sep = "_"),
    date_month = floor_date(date_episode_start, unit = "month")) %>%
  select(
    id,
    episode_id,
    date_month,
    date_episode_start,
    date_admission,
    date_discharge,
    date_episode_end,
    admission_type,
    loc_discharge,
    main_condition,
    other_condition1,
    other_condition2,
    other_condition3,
    other_condition4,
    other_condition5)

#summarise urgency of admission (for each spell)
smr01_long <-
  smr01_long %>%
  mutate(
    admission_type_ndl =
      case_when(admission_type %in% c(10:19) ~ "routine",
                admission_type %in% c(30:39) ~ "emergency",
                admission_type %in% c(20:29) ~ "emergency"))

#select date range and those in NDL population
smr01_long <- 
  filter(smr01_long, 
         date_episode_start >= ymd("2018-03-01"),
         date_episode_start <= ymd("2020-07-31"),
         id %in% unique_ids)
```


# Combine hospital spells
```{r}
#if any spell was an emergency, and emergency admission, else routine
smr01 <-
  smr01_long %>%
  group_by(episode_id) %>%
  arrange(admission_type_ndl) %>%
  summarise(admission_types = paste(admission_type_ndl, collapse = ", ")) %>%
  mutate(admission_type = 
           ifelse(str_detect(admission_types, "emergency"),
                  "emergency", "routine")) %>%
  select(-admission_types)

smr01 <-
  smr01_long %>%
  group_by(episode_id) %>%
  arrange(date_episode_start) %>%
  slice_head() %>%
  select(id,
         episode_id,
         date_month,
         date_episode_start,
         date_episode_end) %>%
  left_join(smr01, by = "episode_id")

#add demog data to hospital admissions
smr01 <-
  smr01 %>%
  left_join(., demog, by = "id") 
```


#Find Covid ICD10 in any diagnosis
```{r}
#make the 5 diagnostic codes per spell into long format
smr01_diagnoses <-
  smr01_long %>%
  pivot_longer(main_condition:other_condition5, values_drop_na = T) %>%
  select(id, episode_id, date_episode_start, name, value)

#pull ids of episodes with either of the 2 Covid ICD10s
covid_episodes <-
  smr01_diagnoses %>%
  filter(str_detect(value, "U071") | str_detect(value, "U072")) %>%
  distinct(episode_id) %>%
  pull(episode_id)

#add flag if episode contains Covid ICD10
smr01 <-
  smr01 %>%
  mutate(covid_episode = ifelse(episode_id %in% covid_episodes, 1, 0))

rm(smr01_diagnoses, covid_episodes)
```

#Find Covid diagnosis in main or other ICD10 code of first spell of episode only
```{r}
#slice off first spell of episode
#make all five ICD10 code fields long
smr01_first_diagnoses <-
  smr01_long %>%
  group_by(episode_id) %>%
  arrange(date_admission) %>%
  slice_head() %>%
  pivot_longer(main_condition:other_condition5, values_drop_na = T) %>%
  select(id, episode_id, date_episode_start, name, value)

#pull ids of episodes with either of the 2 Covid ICD10s
covid_first_episodes <-
  smr01_first_diagnoses %>%
  filter(str_detect(value, "U071") | str_detect(value, "U072")) %>%
  distinct(episode_id) %>%
  pull(episode_id)

#add flag if episode contains Covid ICD10
smr01 <-
  smr01 %>%
  mutate(covid_first_episode = ifelse(episode_id %in% covid_first_episodes, 1, 0)) %>%
  ungroup()

rm(smr01_first_diagnoses, covid_first_episodes)
```

#Covid ICD10s table
```{r}
#table for any diagnosis
#filter pre-March Covid ICD10s as error
covid_any <-
  smr01 %>%
  filter(date_episode_start > ymd("2020-03-01")) %>%
  filter(covid_episode == 1) %>%
  group_by(date_month) %>%
  summarise(people_with_any_covid_diagnosis = n_distinct(id))

#table for first spell diagnosis
#filter pre-March Covid ICD10s as error
covid_first <-
  smr01 %>%
  filter(date_episode_start > ymd("2020-03-01")) %>%
  filter(covid_first_episode == 1) %>%
  group_by(date_month) %>%
  summarise(people_with_covid_diagnosis_first_spell = n_distinct(id),
            cohort_size = n_shielders)

full_join(covid_any, covid_first, by = "date_month") %>%
  write_csv(., "covid_diagnoses.csv")

rm(covid_any, covid_first)
```


#Hospital admissions tables
```{r}
smr01 %>%
  group_by(date_month, .drop = F) %>%
  summarise(hospital_admissions = n(),
            people_admitted = n_distinct(id),
            cohort_size = n_shielders) %>%
  write_csv(., "hospital_admissions_total.csv")

smr01 %>%
  group_by(date_month, age_band, .drop = F) %>%
  summarise(hospital_admissions = n(),
            people_admitted = n_distinct(id),
            cohort_size = n_shielders) %>%
  write_csv(., "hospital_admissions_by_age.csv")

smr01 %>%
  group_by(date_month, sex, .drop = F) %>%
  summarise(hospital_admissions = n(),
            people_admitted = n_distinct(id),
            cohort_size = n_shielders) %>%
  write_csv(., "hospital_admissions_by_sex.csv")

#remove na small numbers
smr01 %>%
  filter(!is.na(simd_quintile)) %>%
  group_by(date_month, simd_quintile, .drop = F) %>%
  summarise(hospital_admissions = n(),
            people_admitted = n_distinct(id),
            cohort_size = n_simd) %>%
  write_csv(., "hospital_admissions_by_deprivation.csv")

smr01 %>%
  group_by(date_month, admission_type, .drop = F) %>%
  summarise(hospital_admissions = n(),
            people_admitted = n_distinct(id),
            cohort_size = n_shielders) %>%
  write_csv(., "hospital_admissions_by_type.csv")

smr01 %>%
  group_by(date_month, admission_type, age_band, .drop = F) %>%
  summarise(hospital_admissions = n(),
            people_admitted = n_distinct(id),
            cohort_size = n_shielders) %>%
  write_csv(., "hospital_admissions_by_type_by_age.csv") 

smr01 %>%
  group_by(date_month, admission_type, sex, .drop = F) %>%
  summarise(hospital_admissions = n(),
            people_admitted = n_distinct(id),
            cohort_size = n_shielders) %>%
  write_csv(., "hospital_admissions_by_type_by_sex.csv") 

#remove NAs due to small numbers
smr01 %>%
  filter(!is.na(simd_quintile)) %>%
  group_by(date_month, admission_type, simd_quintile, .drop = F) %>%
  summarise(hospital_admissions = n(),
            people_admitted = n_distinct(id),
            cohort_size = n_simd) %>%
  write_csv(., "hospital_admissions_by_type_by_deprivation.csv") 
```

#Deaths in Hospital
```{r}
deaths_in_hospital <-
  vroom(here("original_data", "Dash416_Shield20201012_Hospital_Deaths_Release_v1.txt"), delim = "¬") %>%
  clean_names() %>%
  rename(id = dash416_release_uid) %>%
  mutate(smr_dod_discharge_date = as_date(smr_dod_discharge_date)) %>%
  filter(smr_dod_discharge_date >= ymd("2020-03-28")) %>%
  filter(smr_dod_discharge_date <= ymd("2020-07-31")) %>%
  pull(id)

demog <-
  demog %>%
  mutate(died_in_hospital =
           ifelse(id %in% deaths_in_hospital, 1, 0))
```

#Death tables
```{r}
#exclude those without death date

demog %>%
  filter(!is.na(month_death_ndl)) %>%
  group_by(month_death_ndl, .drop = F) %>%
  summarise(n_deaths = n(),
            died_in_hospital = sum(died_in_hospital == 1),
            cohort_size = n_shielders) %>%
  write_csv(., "deaths_total.csv")

demog %>%
  filter(!is.na(month_death_ndl)) %>%
  group_by(month_death_ndl, age_band, .drop = F) %>%
  summarise(n_deaths = n(),
            died_in_hospital = sum(died_in_hospital == 1)) %>%
  write_csv(., "deaths_by_age.csv")

demog %>%
  filter(!is.na(month_death_ndl)) %>%
  group_by(month_death_ndl, sex, .drop = F) %>%
  summarise(n_deaths = n(),
            died_in_hospital = sum(died_in_hospital == 1)) %>%
  write_csv(., "deaths_by_sex.csv")

#NB removed na small numbers
demog %>%
  filter(!is.na(month_death_ndl)) %>%
  filter(!is.na(simd_quintile)) %>%
  group_by(month_death_ndl, simd_quintile, .drop = F) %>%
  summarise(n_deaths = n(),
            died_in_hospital = sum(died_in_hospital == 1)) %>% 
  write_csv(., "deaths_by_deprivation.csv")
```

